Comunicações Hiperligadas: Novo Paradigma de Comunicação e Colaboração, potenciado pela Tecnologia WebRTC

### 2015-05-16 tese/state01.tex
[ADD] goodow, [START] architecture

Initially \ac{NAT} offered an alternative to address exhaustion and a minimal sensation of security. There are four types of \ac{NAT} implementations\cite{rfc3489}, \textit{Full Cone NAT}, \textit{Restricted Cone NAT}, \textit{Port Restricted Cone NAT}, \textit{Symmetric NAT}.

\textit{Full Cone} \ac{NAT} maps public each \ac{IP} address and port to a private \ac{IP} address and port, any external host can communicate with private hosts through their mapped public address and port. This represents the least restrictive type of \ac{NAT} and as we will later, the unique type of \ac{NAT} that enables real time communications from point to point.

\textit{Restricted Cone} \ac{NAT} requires that a private client must first send a message to an external host before it can receive messages from the same host. With this type of \ac{NAT}, the private client can be contacted from any port of the same external host.

\textit{Port Restricted Cone} \ac{NAT} works in the same way as Restricted Cone \ac{NAT}, but it only allows communications from the same external host's IP address and port, ignoring all messages from other applications within the same external host.

\textit{Symmetric} NAT maps different ports for each connection, as we will see later, this type of \ac{NAT} represents a problem on real time communications.

\textit{Asymmetric} \ac{NAT} became a vulgar configuration on the web. As a direct result, problems started to appear, the amount of ports that \ac{IP} makes available is also small compared to our current needs, worse than that, \ac{NAT} also difficult end-to-end communication, forcing most of applications that follow this model to be implemented ineffectively.

@@

On the other hand, \ac{TURN} uses public servers to redirect traffic between private endpoints, it may use a \ac{P2P} network relay to find the best peer, but after that the behavior is much like client-server. Direct communication is only achieved by \ac{STUN} when \ac{NAT} is a type \textit{full cone}. \ac{ICE} uses \ac{STUN} when it's possible and \ac{TURN} otherwise.

@@

\ac{WebRTC}'s \textit{DataChannel} is built on top of \ac{SCTP}, which is encapsulated by \ac{DTLS}. \ac{DTLS} encapsulation provides confidentiality, authentication and integrity over the transfered data. A \textit{Data Channel} has one incoming stream and one outgoing stream, providing bidirectional communication. Each data channel direction can be tweaked for reliable or unreliable transmission, the same can be done for order delivery and priority which can also be defined for improving the quality of service for a particular stream compared to another.

\ac{WebRTC}'s \textit{MediaStream} is built on top of \ac{SRTP}, which requires an external mechanism for key exchange. \ac{DTLS} keys are negotiated on handshake in order to achieve a secure connection. The new keys derived from \ac{DTLS} handshake are seized for \ac{SRTP} encryption, the remaining communications are done through \ac{UDP} without using \ac{DTLS}.

\textit{Skype}\footnote{\url{http://www.skype.com/}} is an application that allows video, voice and instant messaging communication over proprietary protocols, its main strength is the amount of users that are using it nowadays. But compared to \textit{Skype}, \ac{WebRTC} applications don't need to be pre-installed.

\textit{Google Hangouts}\footnote{\url{http://plus.google.com/hangouts}} is a video conference web application, in the past in order to use \textit{hangouts} on a web browser a plug-in was needed to be installed, nowadays hangouts is using \ac{WebRTC}.

\textit{Jitsi Meet}\footnote{\url{http://jitsi.org/Projects/JitsiMeet}} is a \ac{WebRTC} collaborative application that uses \textit{Jitsi Videobridge} for high quality and scalable video conferences and supports shared document editing. \textit{Jitsi Videobridge} is a server that enables multi-party video calls.

>>>>

Initially \ac{NAT} offered an alternative to address exhaustion and a minimal sensation of security. There are four types of \ac{NAT} implementations\cite{rfc3489}, \emph{Full Cone NAT}, \emph{Restricted Cone NAT}, \emph{Port Restricted Cone NAT}, \emph{Symmetric NAT}.

\emph{Full Cone} \ac{NAT} maps public each \ac{IP} address and port to a private \ac{IP} address and port, any external host can communicate with private hosts through their mapped public address and port. This represents the least restrictive type of \ac{NAT} and as we will later, the unique type of \ac{NAT} that enables real time communications from point to point.

\emph{Restricted Cone} \ac{NAT} requires that a private client must first send a message to an external host before it can receive messages from the same host. With this type of \ac{NAT}, the private client can be contacted from any port of the same external host.

\emph{Port Restricted Cone} \ac{NAT} works in the same way as Restricted Cone \ac{NAT}, but it only allows communications from the same external host's IP address and port, ignoring all messages from other applications within the same external host.

\emph{Symmetric} NAT maps different ports for each connection, as we will see later, this type of \ac{NAT} represents a problem on real time communications.

\emph{Asymmetric} \ac{NAT} became a vulgar configuration on the web. As a direct result, problems started to appear, the amount of ports that \ac{IP} makes available is also small compared to our current needs, worse than that, \ac{NAT} also difficult end-to-end communication, forcing most of applications that follow this model to be implemented ineffectively.

@@

On the other hand, \ac{TURN} uses public servers to redirect traffic between private endpoints, it may use a \ac{P2P} network relay to find the best peer, but after that the behavior is much like client-server. Direct communication is only achieved by \ac{STUN} when \ac{NAT} is a type \emph{full cone}. \ac{ICE} uses \ac{STUN} when it's possible and \ac{TURN} otherwise.

@@

\ac{WebRTC}'s \emph{DataChannel} is built on top of \ac{SCTP}, which is encapsulated by \ac{DTLS}. \ac{DTLS} encapsulation provides confidentiality, authentication and integrity over the transfered data. A \emph{Data Channel} has one incoming stream and one outgoing stream, providing bidirectional communication. Each data channel direction can be tweaked for reliable or unreliable transmission, the same can be done for order delivery and priority which can also be defined for improving the quality of service for a particular stream compared to another.

\ac{WebRTC}'s \emph{MediaStream} is built on top of \ac{SRTP}, which requires an external mechanism for key exchange. \ac{DTLS} keys are negotiated on handshake in order to achieve a secure connection. The new keys derived from \ac{DTLS} handshake are seized for \ac{SRTP} encryption, the remaining communications are done through \ac{UDP} without using \ac{DTLS}.

\emph{Skype}\footnote{\url{http://www.skype.com/}} is an application that allows video, voice and instant messaging communication over proprietary protocols, its main strength is the amount of users that are using it nowadays. But compared to \emph{Skype}, \ac{WebRTC} applications don't need to be pre-installed.

\emph{Google Hangouts}\footnote{\url{http://plus.google.com/hangouts}} is a video conference web application, in the past in order to use \emph{hangouts} on a web browser a plug-in was needed to be installed, nowadays hangouts is using \ac{WebRTC}.

\emph{Jitsi Meet}\footnote{\url{http://jitsi.org/Projects/JitsiMeet}} is a \ac{WebRTC} collaborative application that uses \emph{Jitsi Videobridge} for high quality and scalable video conferences and supports shared document editing. \emph{Jitsi Videobridge} is a server that enables multi-party video calls.

### 2015-05-16 tese/state02.tex
[ADD] goodow, [START] architecture

  \ac{WebRTC} doesn't implement signaling, different applications may require different protocols, there is no single answer that fits all problems. Amongst multiple options, signaling can be done by using \ac{SIP}, \ac{XMPP}, \textit{WebSockets}, \textit{Socket.io} or by implementing a custom protocol.

@@

  The WebSocket protocol \cite{rfc6455} provides bidirectional communications over a full-duplex socket channel. WebSocket handshake phase specifies an \ac{HTTP} header in order to upgrade to \textit{WebSocket} type of communication, the remainder messages are done without \ac{HTTP} headers, which leads to much smaller messages and better network usage. WebSockets may not be available on every web browser, frameworks like \textit{socket.io}\footnote{\url{http://socket.io/}} and \textit{SockJS}\footnote{\url{http://github.com/sockjs}} uses \ac{HTTP} when there is no support for WebSockets.

@@

  \ac{SIP} is used in \ac{VoIP} applications due to its compatibility with \ac{PSTN}. Service providers making their \ac{SIP} infrastructures available through WebSockets. Frameworks like \textit{jsSIP}\footnote{\url{http://jssip.net/}}, \textit{QoffeeSIP}\footnote{\url{http://qoffeesip.quobis.com/}} and \textit{sipML5}\footnote{\url{http://sipml5.org/}} are used on client side to parse and encode \ac{SIP} messages, making \ac{SIP} accessible to web based applications.

  \ac{SIP} with \textit{WebSockets} can be used as a signaling method for \ac{WebRTC} applications, it allows web browsers to have audio, video and \ac{SMS} capabilities like mobile phones. For instance, it's possible to inter-operate web communications with \ac{SIP} networks, mobile and fixed phones.

@@

  Amongst multiple XMPP server solutions are: \textit{ejabberd}\footnote{\url{http://jabberd.im/}}, \textit{Metronome}\footnote{\url{http://lightwitch.org/metronome}}, \textit{Openfire}\footnote{\url{http://igniterealtime.org/projects/openfire/}} and \textit{Prosody}\footnote{\url{http://prosody.im/}}. \textit{Ejabberd} is the one that implements more \ac{RFC} specifications and \ac{XEP}s\footnote{\url{http://en.wikipedia.org/wiki/Comparison_of_XMPP_server_software}}.

  A more interesting approach for signaling would be \ac{SigOfly}\cite{sigofly} which allows inter-domain real-time communications. \ac{SigOfly} achieves inter-domain communication by making use of the \textit{Identity Providers} of each peer.

@@

  \ac{SigOfly} is an approach very flexible because participants on a video call are not tied to just one type o signaling implementation. Another important aspect of  \ac{SigOfly} is the ability to perform multi-party conversations either through a \textit{Mesh Topology} or a \textit{Multipoint Control Unit}.

>>>>

  \ac{WebRTC} doesn't implement signaling, different applications may require different protocols, there is no single answer that fits all problems. Amongst multiple options, signaling can be done by using \ac{SIP}, \ac{XMPP}, \emph{WebSockets}, \emph{Socket.io} or by implementing a custom protocol.

@@

  The WebSocket protocol \cite{rfc6455} provides bidirectional communications over a full-duplex socket channel. WebSocket handshake phase specifies an \ac{HTTP} header in order to upgrade to \emph{WebSocket} type of communication, the remainder messages are done without \ac{HTTP} headers, which leads to much smaller messages and better network usage. WebSockets may not be available on every web browser, frameworks like \emph{socket.io}\footnote{\url{http://socket.io/}} and \emph{SockJS}\footnote{\url{http://github.com/sockjs}} uses \ac{HTTP} when there is no support for WebSockets.

@@

  \ac{SIP} is used in \ac{VoIP} applications due to its compatibility with \ac{PSTN}. Service providers making their \ac{SIP} infrastructures available through WebSockets. Frameworks like \emph{jsSIP}\footnote{\url{http://jssip.net/}}, \emph{QoffeeSIP}\footnote{\url{http://qoffeesip.quobis.com/}} and \emph{sipML5}\footnote{\url{http://sipml5.org/}} are used on client side to parse and encode \ac{SIP} messages, making \ac{SIP} accessible to web based applications.

  \ac{SIP} with \emph{WebSockets} can be used as a signaling method for \ac{WebRTC} applications, it allows web browsers to have audio, video and \ac{SMS} capabilities like mobile phones. For instance, it's possible to inter-operate web communications with \ac{SIP} networks, mobile and fixed phones.

@@

  Amongst multiple XMPP server solutions are: \emph{ejabberd}\footnote{\url{http://jabberd.im/}}, \emph{Metronome}\footnote{\url{http://lightwitch.org/metronome}}, \emph{Openfire}\footnote{\url{http://igniterealtime.org/projects/openfire/}} and \emph{Prosody}\footnote{\url{http://prosody.im/}}. \emph{Ejabberd} is the one that implements more \ac{RFC} specifications and \ac{XEP}s\footnote{\url{http://en.wikipedia.org/wiki/Comparison_of_XMPP_server_software}}.

  A more interesting approach for signaling would be \ac{SigOfly}\cite{sigofly} which allows inter-domain real-time communications. \ac{SigOfly} achieves inter-domain communication by making use of the \emph{Identity Providers} of each peer.

@@

  \ac{SigOfly} is an approach very flexible because participants on a video call are not tied to just one type o signaling implementation. Another important aspect of  \ac{SigOfly} is the ability to perform multi-party conversations either through a \emph{Mesh Topology} or a \emph{Multipoint Control Unit}.

### 2015-05-16 tese/state03.tex
[ADD] goodow, [START] architecture

  Hyper-video is a kind of video that contains links to any kind of hypermedia, including links to skip part of it. An example of hypermedia application could be a search engine over hypermedia content, like subtitles, in order to jump to a specific time. \textit{HyperCafe} \cite{hypercafe} was an experimental project to expose hyper-video concepts that consisted on an interactive film by switching between conversations inside a cafe.

  Detail-on-demand is a subset of hyper-video that allow us to obtain additional information about something that appears along the video, like obtaining information about a painting that appears in a particular segment. \textit{Hyper-Hitchcock}\cite{hitchcock} is an editor and player of detail-on-demand video.

@@

  \textit{HyVAL}\cite{hyval} is an \ac{XML} based language that was proposed for modeling composition, synchronization and interaction of hypermedia. HyVAL defines defines video structure, internal video and external media objects.

  \textit{HyVAL}'s video structure object defines a structure derived from traditional video, which divides video into segments, scenes, shots and frames hierarchically, this approach is quite restrictive if we want to apply hyper-video concepts to videos that don't follow this structure. External media objects are linked by primary video, those objects can represent other videos, images, text, animation and sound.

@@

  \ac{SMIL} is an \ac{XML} based language that defines twelve modules: \textit{Animation}, \textit{Content Control}, \textit{Layout}, \textit{Linking}, \textit{Media Objects}, \textit{SmilText}, \textit{Metainformation}, \textit{Structure}, \textit{Timing}, \textit{Time Manipulations}, \textit{State} and \textit{Transitions}.

@@

  \item \textbf{Structure} module defines the basic elements and attributes for structuring \ac{SMIL} content. This module defines a \textit{head} element that contains non temporal behavior information defined by  \textit{Metainformation}, \textit{Layout} and \textit{Content Control} modules. This module also defines defines the \textit{body} element, where all temporal related module information is contained.

  \item \textbf{Timing} module is the most important module on \ac{SMIL} specification, due to its complexity, it is divided into seventeen sub-modules for coordination and synchronization of media over time. The three main elements are \textit{seq}, \textit{excl} and \textit{par}, respectively, they play child elements in sequence, one at a time and all at the same time.

@@

  \item \textbf{Transitions} module defines attributes and elements for transitions across multiple \ac{SMIL} elements according to \textit{Timing} module.

@@

Like \ac{DOM}, \ac{SMIL} \ac{DOM} is an \ac{API} for \ac{SMIL} documents. Allowing \ac{CRUD} operations over \ac{SMIL} documents is an important feature for extending \ac{SMIL} capabilities, for example for creating non-linear animations and triggering external events like \textit{JavaScript} functions.

@@

  \ac{HTML} is a markup language based on \ac{XML} that is used for creating web pages. \ac{HTML} alone is a very poor language when we are focused on visual appealing and interactive web pages. Languages like \ac{CSS} and \textit{JavaScript} are typically combined to \ac{HTML} for improving the interaction and appearance of a web page.

@@

  \textit{JavaScript} is an imperative object-oriented language based on \textit{ECMAScript}. It is used mainly on client-side and executed by a web browser. \textit{JavaScript} has its own implementation of \ac{DOM} and one of its advantages is the ability to download and execute code on the fly without the need of pre-installed plug-ins.

  \textit{JavaScript} has compatibility issues amongst different web browsers, leading to different behaviors. To solve that problem, there are libraries written in \textit{JavasSript}, namely \textit{jQuery} which implements the same functionality for multiple browsers, solving most of incompatibility issues.

  SmillingWeb \cite{smillingweb} attempts to implement a cross platform multimedia player designed for \ac{SMIL} 3.0 presentations with \textit{JavaScript} and \textit{jQuery} which, unlike \cite{ambulant}, doesn't need to be installed and shouldn't have incompatibility issues.

  SmillingWeb already takes advantage of \ac{HTML}5 and \ac{CSS}3, they take into account unsupported web browsers though \textit{Modernizr}\footnote{\url{http://modernizr.com/}}, a simple \textit{JavaScript} library that may require plug-ins if new features aren't supported.

@@

  With the emergence of \ac{HTML}5, tags like \textit{video}, \textit{audio} and \textit{track} allow us to play video with multiple codecs, audio and subtitles in \ac{WebVTT} format. Another important tag is \textit{canvas} that allows to draw graphics with \textit{JavaScript} on a rectangle within a web page.

  \ac{SVG} is an \ac{XML} based format that incorporates the animation module of \ac{SMIL}. Currently, \ac{SVG} allows to add movement and animate attributes of elements. When embedded on \ac{HTML}5, it allows dynamic changes to inner content in real-time through \ac{DOM} \ac{API}, besides that, it also allows to call \textit{JavaScript} functions on events such as animation end, mouse over and mouse click.

  Video and audio features are already possible with \ac{HTML}5, like \ac{SVG} it is also possible to bind \textit{JavaScript} functions for different kinds of events over video and audio elements.

  Back in 1995, \textit{flash}\footnote{\url{http://www.adobe.com/products/flash.html}} was developed for web-based animations, introducing video support in 2002, \textit{flash} started to grow after that. Concurrency players, at that time, were focused on playing video and audio, \textit{flash} had vector graphics and they were focused on streaming \textit{on-demand} video across multiple platforms. \textit{VP6} was their choice on video codecs, optimizing for half of video size for the same quality and providing adjusted video quality based on \textit{Internet} connection latency.

  \textit{Adobe Flash} was the most widely used applications for reproducing live broadcast and recorded video \cite{flashvideo}, it supports progressive video download using \ac{HTTP} and streaming using \ac{RTMP}.

@@

  Although \ac{HTML}5, \textit{JavaScript}, \ac{CSS} and \ac{WebRTC} are implementing some features of flash, it doesn't mean that flash will be replaced, instead of that both technologies can be used to develop rich \textit{Internet} applications. It is also important to note that \ac{HTML}5 is more compatible with mobile devices than adobe flash.

  Like \textit{Flash}, \textit{Microsoft Silverlight}\footnote{\url{http://www.microsoft.com/silverlight/}} is a cross browser plug-in and platform that is used to develop rich \textit{Internet} applications. It supports vector graphics, animation and video. Compared to flash, which uses \textit{ActionScript}, \textit{Silverlight} applications can use languages like C\#, \textit{VisualBasic} and \ac{XAML}. \textit{Silverlight} uses a technique called \textit{Smooth Streaming} from \textit{IIS Media Service} that consists on delivering video in real-time with adjusted quality in function of bandwidth changing and \ac{CPU} usage.

  Using technologies that relies only on web standards, like \ac{CSS}, \ac{HTML}5, \textit{JavaScript} and \ac{SVG}, will make possible to raise communications to a new level. For example, with \ac{API}s like WebGL\footnote{\url{http://khronos.org/webgl/}}, it is now possible to manipulate a three dimensional environment in the context of a hyper-call. Another example would be a collaborative spreadsheet using \ac{WebRTC}. With this, hyper-calls are not limited to only audio, image, text and video, but also interaction with complex graphical user interfaces that changes over time.

>>>>

  Hyper-video is a kind of video that contains links to any kind of hypermedia, including links to skip part of it. An example of hypermedia application could be a search engine over hypermedia content, like subtitles, in order to jump to a specific time. \emph{HyperCafe} \cite{hypercafe} was an experimental project to expose hyper-video concepts that consisted on an interactive film by switching between conversations inside a cafe.

  Detail-on-demand is a subset of hyper-video that allow us to obtain additional information about something that appears along the video, like obtaining information about a painting that appears in a particular segment. \emph{Hyper-Hitchcock}\cite{hitchcock} is an editor and player of detail-on-demand video.

@@

  \emph{HyVAL}\cite{hyval} is an \ac{XML} based language that was proposed for modeling composition, synchronization and interaction of hypermedia. HyVAL defines defines video structure, internal video and external media objects.

  \emph{HyVAL}'s video structure object defines a structure derived from traditional video, which divides video into segments, scenes, shots and frames hierarchically, this approach is quite restrictive if we want to apply hyper-video concepts to videos that don't follow this structure. External media objects are linked by primary video, those objects can represent other videos, images, text, animation and sound.

@@

  \ac{SMIL} is an \ac{XML} based language that defines twelve modules: \emph{Animation}, \emph{Content Control}, \emph{Layout}, \emph{Linking}, \emph{Media Objects}, \emph{SmilText}, \emph{Metainformation}, \emph{Structure}, \emph{Timing}, \emph{Time Manipulations}, \emph{State} and \emph{Transitions}.

@@

  \item \textbf{Structure} module defines the basic elements and attributes for structuring \ac{SMIL} content. This module defines a \emph{head} element that contains non temporal behavior information defined by  \emph{Metainformation}, \emph{Layout} and \emph{Content Control} modules. This module also defines defines the \emph{body} element, where all temporal related module information is contained.

  \item \textbf{Timing} module is the most important module on \ac{SMIL} specification, due to its complexity, it is divided into seventeen sub-modules for coordination and synchronization of media over time. The three main elements are \emph{seq}, \emph{excl} and \emph{par}, respectively, they play child elements in sequence, one at a time and all at the same time.

@@

  \item \textbf{Transitions} module defines attributes and elements for transitions across multiple \ac{SMIL} elements according to \emph{Timing} module.

@@

Like \ac{DOM}, \ac{SMIL} \ac{DOM} is an \ac{API} for \ac{SMIL} documents. Allowing \ac{CRUD} operations over \ac{SMIL} documents is an important feature for extending \ac{SMIL} capabilities, for example for creating non-linear animations and triggering external events like \emph{JavaScript} functions.

@@

  \ac{HTML} is a markup language based on \ac{XML} that is used for creating web pages. \ac{HTML} alone is a very poor language when we are focused on visual appealing and interactive web pages. Languages like \ac{CSS} and \emph{JavaScript} are typically combined to \ac{HTML} for improving the interaction and appearance of a web page.

@@

  \emph{JavaScript} is an imperative object-oriented language based on \emph{ECMAScript}. It is used mainly on client-side and executed by a web browser. \emph{JavaScript} has its own implementation of \ac{DOM} and one of its advantages is the ability to download and execute code on the fly without the need of pre-installed plug-ins.

  \emph{JavaScript} has compatibility issues amongst different web browsers, leading to different behaviors. To solve that problem, there are libraries written in \emph{JavasSript}, namely \emph{jQuery} which implements the same functionality for multiple browsers, solving most of incompatibility issues.

  SmillingWeb \cite{smillingweb} attempts to implement a cross platform multimedia player designed for \ac{SMIL} 3.0 presentations with \emph{JavaScript} and \emph{jQuery} which, unlike \cite{ambulant}, doesn't need to be installed and shouldn't have incompatibility issues.

  SmillingWeb already takes advantage of \ac{HTML}5 and \ac{CSS}3, they take into account unsupported web browsers though \emph{Modernizr}\footnote{\url{http://modernizr.com/}}, a simple \emph{JavaScript} library that may require plug-ins if new features aren't supported.

@@

  With the emergence of \ac{HTML}5, tags like \emph{video}, \emph{audio} and \emph{track} allow us to play video with multiple codecs, audio and subtitles in \ac{WebVTT} format. Another important tag is \emph{canvas} that allows to draw graphics with \emph{JavaScript} on a rectangle within a web page.

  \ac{SVG} is an \ac{XML} based format that incorporates the animation module of \ac{SMIL}. Currently, \ac{SVG} allows to add movement and animate attributes of elements. When embedded on \ac{HTML}5, it allows dynamic changes to inner content in real-time through \ac{DOM} \ac{API}, besides that, it also allows to call \emph{JavaScript} functions on events such as animation end, mouse over and mouse click.

  Video and audio features are already possible with \ac{HTML}5, like \ac{SVG} it is also possible to bind \emph{JavaScript} functions for different kinds of events over video and audio elements.

  Back in 1995, \emph{flash}\footnote{\url{http://www.adobe.com/products/flash.html}} was developed for web-based animations, introducing video support in 2002, \emph{flash} started to grow after that. Concurrency players, at that time, were focused on playing video and audio, \emph{flash} had vector graphics and they were focused on streaming \emph{on-demand} video across multiple platforms. \emph{VP6} was their choice on video codecs, optimizing for half of video size for the same quality and providing adjusted video quality based on \emph{Internet} connection latency.

  \emph{Adobe Flash} was the most widely used applications for reproducing live broadcast and recorded video \cite{flashvideo}, it supports progressive video download using \ac{HTTP} and streaming using \ac{RTMP}.

@@

  Although \ac{HTML}5, \emph{JavaScript}, \ac{CSS} and \ac{WebRTC} are implementing some features of flash, it doesn't mean that flash will be replaced, instead of that both technologies can be used to develop rich \emph{Internet} applications. It is also important to note that \ac{HTML}5 is more compatible with mobile devices than adobe flash.

  Like \emph{Flash}, \emph{Microsoft Silverlight}\footnote{\url{http://www.microsoft.com/silverlight/}} is a cross browser plug-in and platform that is used to develop rich \emph{Internet} applications. It supports vector graphics, animation and video. Compared to flash, which uses \emph{ActionScript}, \emph{Silverlight} applications can use languages like C\#, \emph{VisualBasic} and \ac{XAML}. \emph{Silverlight} uses a technique called \emph{Smooth Streaming} from \emph{IIS Media Service} that consists on delivering video in real-time with adjusted quality in function of bandwidth changing and \ac{CPU} usage.

  Using technologies that relies only on web standards, like \ac{CSS}, \ac{HTML}5, \emph{JavaScript} and \ac{SVG}, will make possible to raise communications to a new level. For example, with \ac{API}s like WebGL\footnote{\url{http://khronos.org/webgl/}}, it is now possible to manipulate a three dimensional environment in the context of a hyper-call. Another example would be a collaborative spreadsheet using \ac{WebRTC}. With this, hyper-calls are not limited to only audio, image, text and video, but also interaction with complex graphical user interfaces that changes over time.

### 2015-05-16 tese/state04.tex
[ADD] goodow, [START] architecture

	{\color{blue} Goodow [12], is a recent FLOSS framework copying the Google Drive Real-Time API with additional clients for Android and iOS, while providing its own server implementation.}

>>>>

	\emph{Goodow}\footnote{\url{http://realtimeplayground.goodow.com/}} is a collaborative framework with its own server implementation, it supports four types of collaborative elements: String, Lists, Maps and Custom objects.
